---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data  
The data was provided in a csv file, with three columns representing: the number of steps (steps), the date (date) and a label for the 5 minute interval (interval). Missing values in steps where coded as NA and dates were coded in the ‘YYYY-MM-DD’ format.

```{r}
setwd("/Users/mac1/Desktop/DataCours")
activity= read.csv("activity.csv", header=TRUE, sep=',')
```

## What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day
```{r}
StepByDay <-aggregate(list(meanSteps=activity$steps), list(date=activity$date), sum, na.rm=TRUE)
hist(StepByDay$meanSteps, breaks = 30, col = "red",border = "black", main="Mean total number of steps taken per day", xlab ="Total steps per day", ylab ="Number of day with this total steps area", xlim=c(0,25000))
```

Calculate mean and median total number of steps taken per day
```{r}
round(mean(StepByDay$meanSteps))
round(median(StepByDay$meanSteps))
```

## What is the average daily activity pattern?

Make a time series plot
```{r}
StepsByInterval <- aggregate(list(meanSteps=activity$steps), list(interval=activity$interval), mean, na.rm=TRUE)
plot(StepsByInterval$meanSteps ~ StepsByInterval$interval  , col="black", main=" Average daily activity pattern", xlab='Interval during a day',  ylab="Mean of number of steps", type='l')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
StepsByInterval$interval[which.max(StepsByInterval$meanSteps)]
```


## Imputing missing values

Calculate and report the total number of missing values in the dataset  
The provided data contains a total of 2304, from which 2304 are missing number of steps, and none are missing dates or interval labels (i.e. all missing values are listed explicitly). As described in the PA1 instructions, the absence of these values might have an impact on the activity estimations; follows a description of the strategy I propose to fill in all the missing values in the dataset.

```{r}
na_steps<-sum(is.na(activity$steps))
na_date<-sum(is.na(activity$date))
na_interval<-sum(is.na(activity$interval))
```

Devise a strategy for filling in all of the missing values in the dataset:    
During 61 days, corresponding to a significant sample, we can expect a repeat of values for a same intervals. 
The nearest behavior for an interval is that which corresponds to the average of the same interval for the other days.
So, we will fill the missing values in the dataset with  the mean for that 5-minute interval

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newactivity <- activity
l=length(newactivity[,1])
for(i in 1:l)
  { 
  if(is.na(newactivity$steps[i]))
    {
    newactivity$steps[i]<-StepsByInterval$meanSteps[StepsByInterval$interval==newactivity$interval[i]]
    }
  }
```

Make a histogram of the total number of steps taken each day
```{r}
NewStepByDay <-aggregate(list(meanSteps=newactivity$steps), list(date=newactivity$date), sum, na.rm=TRUE)
hist(NewStepByDay$meanSteps, breaks = 30, col = "red",border = "black", main="Mean total number of steps taken per day", xlab ="Total steps per day", ylab ="Number of day with this total steps area", xlim=c(0,25000))
```

Calculate and report the mean and median total number of steps taken per day.

```{r}
round(mean(NewStepByDay$meanSteps))
round(median(NewStepByDay$meanSteps))
```

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 

```{r}
newactivity$dateType<-as.POSIXlt(newactivity$date)$wday
l=length(newactivity[,1])
for(i in 1:l)
{ 
  if(newactivity$dateType[i]==0)
  {
    newactivity$weekend[i]<-"weekend"
  }
  
  if(newactivity$dateType[i]==6)
  {
    newactivity$weekend[i]<-"weekend"
  }
  
  else
  {
    newactivity$weekend[i]<-"weekday"
  }
  
}
```

Make a panel plot containing a time series plot

```{r}
StepsByWeekEnd <- aggregate(newactivity$steps ~ newactivity$weekend + newactivity$interval,newactivity, mean)
with(StepsByWeekEnd, {
  plot(StepsByWeekEnd[,3] ~ StepsByWeekEnd[,2], subset = StepsByWeekEnd[,1]== "weekend" , col="blue", main=" Average daily activity pattern", xlab='Interval during a day',  ylab="Mean of number of steps", type='l')
  legend("topright",  lty=1, col = c("red","blue"), legend = c("weekday activity", "weekend activity"), bty='n')
  lines(StepsByWeekEnd[,3] ~ StepsByWeekEnd[,2], subset = StepsByWeekEnd[,1]== "weekday" , col="red",  type='l')
   })
```
